Space is the Place: Effects of Continuous Spatial Structure on Analysis of Population Genetic Data

### 2019-12-05 ms/review-responses.tex

\begin{point}{\revref}
    Perhaps 'Demographic Inference' might better reflect the content of this section.
\end{point}

\reply{
Good suggestion -- we have changed the section heading to ``demographic inference'' \revref.
}

\begin{point}{\revref}
    This sentence with 'Gaussian noise with mean zero and standard deviation 10' is confusing since it was mentioned earlier that the modelled phenotype must vary as human height across Europe, and human height varies 2 standard deviations. Only after reading the whole paragraph it becomes clear that 'standard deviation 10' here referes to unit of height. Please consider rephrasing this sentence.
\end{point}

\reply{
We have revised this sentence to clarify that we aim to produce a variation in mean phenotype of two standard deviations across the landscape \revref.
}

@@

    We definately think that a comparison to the spatial Lambda-Fleming-Viot
    would be an extremely useful thing to do, for the reasons the reviewer mentions,
    but this is beyond the scope of the current study.
    The SLFV model is only known to be an approximation of one particular biologically explicit model
    (of patchy extinction-recolonization)
    and is only conjectured to be a good approximation
    for spatial models more generally.
    To reiterate, the definition of the SLFV was explicitly motivated by a desire to approximate
    spatial models, especially Bolker-Pacala models like ours.
    Therefore, our study does important preliminary work
    for simulation study of the spatial Lambda-Fleming-Viot,
    since it works through some of the issues in simulating biologically realistic spatial models,
    and describes some of the patterns.

\begin{point}{\autoref{fig:genparams}:}
    I do not understand why the neighborhood size varies to the same extent in the random mating model as it does for the spatial model. For the random mating model, I would have expected the neighborhood size to be equal to the census size since all individuals have the same probability of being a parent of any given offspring. From [the paragraph at \revref], it is clear that the spatial model would converge to the random mating model when the mean parent-offspring distance tends to infinity only if we were to ignore the impact of range edges. I am thus wondering whether the variation of neighborhood size one observes in Fig 2 for the random mating model is just a consequence of border effects. If that is the case, then the authors should state it clearly and try to justify it from a biological perspective.
\end{point}

\reply{
    It's a good point that neighborhood size, as defined to be ``number of possible parents'',
    should indeed be the census size.
    However, we are using $N_W$ defined in the same way for both spatial and random mating simulations
    as a way of comparing spatial to ``nonspatial'' models
    keeping other aspects of demography (almost) the same.
    This is indeed part of what's going on.
    We have added some discussion of the census size scaling at \revref{} and \llname{censusscaling},
    but haven't added a detailed discussion of whether or not most species experience
    decreased habitat quality at range edges,
    since that seems like it is getting too far into the weeds.
}

@@

\begin{point}{}
    GWAS mixed models. To what extent can spatial signals (e.g. corner, patchy) be corrected with mixed models, e.g. with PCs and PC-adjusted GRM as in Conomos et al, 2016 using PC-AiR and PC-Relate)? Is patchiness related to dispersal? I'm curious how this relates to the predictive ability of GWAS phenotypes with some spatial association that may or may not be associated with environmental effects.
\end{point}

\reply{
    Good question -- we are also interested to know how mixed models perform here, but think that to properly test that we would want to change our design to generate phenotypes from simulated genotypes. This would allow us to evaluate false-negatives in addition to false-positives. This is important because, if mixed models do provide stronger control for stratification they are also likely to remove true signals of causal SNPs if those SNPs covary with spatial structure. We now point to these methods explicitly in the discussion \llname{mixedmodelcallout}, but think that incorporating that study here would make this paper too long and we plan to explore this timely issue in the future. We also think the PC results are still quite relevant as the method is quite population in current studies. 
}

\begin{point}{}
    Code availability. This github link doesn't work, but is important to be able to evaluate for review: \url{https://github.com/petrelharp/spaceness}.
\end{point}

\reply{
Apologies, it was accidentally set to private. The link should work now
(and has been replaced by the primary link, \url{https://github.com/kern-lab/spaceness}).
}
